Red Wine Quality analysis by Lazzat Sultanbek
========================================================

```{r echo=FALSE,warning=FALSE,message=FALSE, global_options}
knitr::opts_chunk$set( echo=FALSE,warning=FALSE,message=FALSE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

knitr::opts_chunk$set(fig.width=9,fig.height=5,fig.path='Figs/',
                      fig.align='center',tidy=TRUE,
                      echo=FALSE,warning=FALSE,message=FALSE)

# load the ggplot graphics package and the others
library(dplyr)
library(ggplot2)
library(ggplot2)
library(GGally)
library(scales)
library(memisc)
library(MASS)
```


```{r echo=FALSE, Load_the_Data}
# Load the Data
redwine <- read.csv("wineQualityReds.csv")

```

The report explores data set on the chemical properties of the wine.

# Univariate Plots Section

```{r echo=FALSE, Data_glimpse}
#Lets take a look at sample dataset
glimpse(redwine)

```

```{r echo=FALSE, Data_Summary}
#The following code is run to see the data summary for each individual variable
summary(redwine)
```

There are 1,599 observations and 12 variables. Note that X represents the 
numbering of the sample and not the variable by itself. Quality is the output 
variable that we will be exploring. 


```{r}
# Transform quality from an integer to an ordered factor
redwine$quality <- factor(redwine$quality, ordered = T)
```

Lets drop unused variable X.

```{r}
# Drop variable X
redwine <- dplyr::select(redwine,-c(X))
```

I am also creating a new varible called rating which will split wine into 3
categories: bad, good and excellent.

```{r}
# Create variable named 'rating'
redwine$rating <- ifelse(redwine$quality < 5, 'bad', ifelse(
  redwine$quality < 7, 'good', 'excellent'))

# Order newly created variable
redwine$rating <- ordered(redwine$rating,
                     levels = c('bad', 'good', 'excellent'))
```

Prior to proceding to plots we cleaned up the data by 1) transforming quality  
from integer to ordered factor, 2) dropping variable X which is just a 
numbering of the data and 3) creating a new variable called rating (bad, 
average, excellent). 


```{r echo=FALSE, Quality_plot}
#Create plot for quality variable
qplot(data = redwine, x = quality, fill = I('#099DD9')) + 
  ggtitle('Quality')

```

Majority of red wine sample got rating of 5 and 6. There are no observations 
were red wine got 1,2 rating or 9,10 rating. 

```{r echo=FALSE, Rating_plot}
#Create plot for variable rating
qplot(data = redwine, x = rating, fill = I('#099DD9')) + 
  ggtitle('Rating')

```

When we plot data by rating we can see that majority of wine falls into "good" 
wine category with quality score of 5-6 and next up comes excellent category 
(scores equal to and higher than 7). Only a small proportion of wine falls into 
"bad" rating category. 


```{r echo=FALSE, Fixed.acidity_Plot}
#Create plot for fixed.acidity
qplot(data = redwine, x = fixed.acidity, binwidth = 0.2, fill = I('#F7A88D')) + 
  ggtitle('Fixed Acidity Plot')

```

```{r}
#Get summary for variable fixed.acidity
summary(redwine$fixed.acidity)

```


Most red wines have fixed acidity ranging between 7 to 9. There are some 
outliers which range up to 15.90. Median value of red wines is 7.90.  

```{r echo=FALSE, Volatile.acidity_Plot}
#Create plot for volatile.acidity
qplot(data = redwine, x = volatile.acidity, binwidth = 0.05, 
      fill = I('#118478')) + 
  ggtitle('Volatile Acidity Plot')

```

```{r}
#Get summary for variable volatile.acidity
summary(redwine$volatile.acidity)
```

Most red wines have volatile acidity ranging between 0.39 to 0.64. There are 
some outliers which range up to 1.58. 


```{r echo=FALSE, Citric.acid_Plot}
#Create plot for citric.acid
qplot(data = redwine, x = citric.acid, binwidth = 0.01, fill = I('#F4B8AB')) + 
  ggtitle('Citric Acid Plot')

```

We see that the data is skewed to the right, lets remove outliers. 

```{r echo=FALSE, message=FALSE, warning=FALSE, Citric.acid_zoomed_in}
#lets zoom in to citric.acid plot by removing outliers

qplot(data = redwine, 
      x = citric.acid, 
      xlim = c(0,0.7), 
      binwidth = 0.03, 
      fill = I('#F4B8AB')) + 
  ggtitle('Citric Acid - zoomed in')

```

```{r echo=FALSE, citric_acid_summary}
#Get summary for variable citric.acid
summary(redwine$citric.acid)
```

Citric acid values range from 0 to 1, although 75% of the population falls 
under 0.42. 


```{r echo=FALSE, Residual_Sugar_plot}
#Create plot for residual.sugar
qplot(data = redwine, x = residual.sugar, binwidth = 0.2, fill = I('#0BA28E')) + 
  ggtitle('Residual Sugar')

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Residual.sugar_zoomed_in}
#lets zoom in to residual.sugar plot by removing outliers
qplot(data = redwine, x = residual.sugar, xlim = c(0,4), binwidth = 0.2, 
      fill = I('#0BA28E')) + 
  ggtitle('Residual Sugar - zoomed in')

```

Residual sugar values tends to vary significantly thoughout the sample 
population. It would be interesting to see if there is any correlation with 
quality of wine. Residual sugar values range from 0 to 16, however zooming in 
we see that most red wines have residual sugar of 1.5 to 2.5. 


```{r echo=FALSE, Chlorides_Plot}
#Create plot for chlorides
qplot(data = redwine, x = chlorides, binwidth = 0.02, fill = I('#0EBFDF')) + 
  ggtitle('Chlorides')

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Chlorides_zoomed_in_Plot}
#lets zoom in to chlorides plot by removing outliers
qplot(data = redwine, x = chlorides, xlim = c(0.01,0.15), binwidth = 0.005, 
      fill = I('#0EBFDF')) + 
  ggtitle('Chlorides zoomed in')

```

Another chemical property skewed to the right is chlorides. Eliminating those 
outliers and zooming in we can see that most red wines have chlorides with 
values of 0.05 to 0.09.  


```{r echo=FALSE, message=FALSE, warning=FALSE, Free_sulfur_dioxide_plot}
#Create plot for free.sulfur.dioxide
qplot(data = redwine, x = free.sulfur.dioxide, binwidth = 2, 
      fill = I('#F59B46')) + 
  ggtitle('Free sulfur dioxide')

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Free_sulfur_dioxide_zoomed_in}
#lets zoom in to free.sulfur.dioxide plot by removing outliers
qplot(data = redwine, x = free.sulfur.dioxide, xlim = c(0,20), binwidth = 0.5, 
      fill = I('#F59B46')) + 
  ggtitle('Free sulfur dioxide - zoomed in')

```


Similar to chlorides and residual sugar, free sulfur dioxide levels also skewed 
to the right. There seems to be a pattern here and it would be interesting to 
later analyze if there is a correlation between these 3 variables. Are the same
samples appearing as outliers in plots for all three (chlorides, residual sugar,
free sulfur dioxide)? 

Zooming into free sulfur dioxide we see that majority of wine range from 3 to 
40.There is a spike at values 5 and 6, which will be interesting to explore 
later. 


```{r echo=FALSE, message=FALSE, warning=FALSE, Total_Sulfur_Dioxide_Plot}
#Create plot for total.sulfur.dioxide
qplot(data = redwine, x = total.sulfur.dioxide, binwidth = 2, 
      fill = I('#17C4B9')) + 
  ggtitle('Total Sulfur Dioxide')

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Total_Sulfur_Dioxide_zoomed_in}
#lets zoom in to total.sulfur.dioxide plot by removing outliers
qplot(data = redwine, x = total.sulfur.dioxide, xlim = c(0,100), 
      binwidth = 2, fill = I('#17C4B9')) + 
  ggtitle('Total Sulfur Dioxide zoomed in')

```

Most red wine in the sample have total sulfur dioxide that ranges from 10 to 50. 


```{r echo=FALSE, Density_Plot}
#Create plot for Density
qplot(data = redwine, x = density, binwidth = 0.001, fill = I('#F4B8AB')) + 
  ggtitle('Density')

```


```{r echo=FALSE, pH_Plots}
#Create plot for pH
qplot(data = redwine, x = pH, binwidth = 0.1, fill = I('#FABCC7')) + 
  ggtitle('pH')

```


Density and pH values look normally distributed, most wines having density level 
from 0.995 to 1 and pH ranging from 3.1 to 3.5. 


```{r echo=FALSE, Sulphates_Plot}
#Create plot for sulphates
qplot(data = redwine, x = sulphates, binwidth = 0.1, fill = I('#FEA679')) + 
  ggtitle('Sulphates')

```

Sulphates are skewed to the right with most wines ranging between 0.5 and 0.7. 

```{r echo=FALSE, Alcohol_Plot}
#Create plot for alcohol
qplot(data = redwine, x = alcohol, binwidth = 0.1, fill = I('#7BCBCA')) + 
  ggtitle('Alcohol')

```

```{r}
#Get summary for variable alcohol
summary(redwine$alcohol)
```

Alcohol level of wines in population has a mean of 10.42 and median of 10.20. 
About 75% of wine have alcohol level lower than 11.10. 


```{r}
#Lets subset the data to see what are the qualities of "excellent" wines
summary(subset(redwine, rating == 'excellent'))
```

```{r}
#Lets subset the data to see what are the qualities of "bad" wines
summary(subset(redwine, rating == 'bad'))
```

```{r}
#Lets subset the data to see what are the qualities of "good" wines
summary(subset(redwine, rating == 'good'))
```

# Univariate Analysis


### What is the structure of your dataset?

There are 1,599 observations in the dataset with 12 variables. At least 3 wine
experts rated the quality of each wine, providing a rating between 0 (very bad)
and 10 (very excellent). Variables in the dataset are:

1 - fixed acidity: most acids involved with wine or fixed or nonvolatile (do 
not evaporate readily).\
2 - volatile acidity: the amount of acetic acid in wine, which at too high of 
levels can lead to an unpleasant, vinegar taste.\
3 - citric acid: found in small quantities, citric acid can add 'freshness' and
flavor to wines.\
4 - residual sugar: the amount of sugar remaining after fermentation stops, 
it's rare to find wines with less than 1 gram/liter and wines with greater than 
45 grams/liter are considered sweet.\
5 - chlorides: the amount of salt in the wine.\
6 - free sulfur dioxide: the free form of SO2 exists in equilibrium between 
molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial 
growth and the oxidation of wine.\
7 - total sulfur dioxide: amount of free and bound forms of S02; in low 
concentrations, SO2 is mostly undetectable in wine, but at free SO2 
concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine.\
8 - density: the density of water is close to that of water depending on the 
percent alcohol and sugar content.\
9 - pH: describes how acidic or basic a wine is on a scale from 0 (very acidic) 
to 14 (very basic); most wines are between 3-4 on the pH scale.\
10 - sulphates: a wine additive which can contribute to sulfur dioxide gas (S02) 
levels, wich acts as an antimicrobial and antioxidant.\
11 - alcohol: the percent alcohol content of the wine.\
\
Output variable (based on sensory data): \
12 - quality (score between 0 and 10)\


### What is/are the main feature(s) of interest in your dataset?

Main feature of interest in the dataset is the quality of wine and other 
variables which directly or in collaboration with other characteristics impact 
the quality. 

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

Comparing qualities of bad, good and excellent wines, volatile.acidity and 
citric.acid differs significantly for each rating and therefore hinting those 
are the qualities impacting the quality of wine. Another characteristic that 
would be interesting to explore is the level of alcohol and its impact on 
quality of wine. 

### Did you create any new variables from existing variables in the dataset?

I created a rating variable to group the population into 3 categories: bad, 
good and excellent. 


### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form 
of the data? If so, why did you do this?

Chlorides, residual sugar and free sulfur dioxide levels are all skewed to the 
right. There seems to be a pattern here and it would be interesting to later 
analyze if there is a correlation between these 3 variables. Are the same 
samples appearing as outliers in plots for all three (chlorides, residual sugar, 
free sulfur dioxide)? 

I deleted the variable X from the dataset which was just a numbering of sample 
and not the characteristic of wine. I also transformed quality from an integer 
to an ordered factor. 


# Bivariate Plots Section

```{r}
#Lets create a new dataset without rating
redwine2 <- dplyr::select(redwine,-c(rating))
redwine2$quality <- as.numeric(redwine2$quality)
round(cor(redwine2),2)

```

Before I start bivariate plots analysis I would like to run ggpairs to see the 
relationship between different variables. 

```{r echo=FALSE, message=FALSE, warning=FALSE, ggpairs, fig.width=9, fig.height=8}
#run ggairs to see relationship between different variables
ggpairs(dplyr::select(redwine,-c(sulphates, residual.sugar,chlorides, pH,
                          total.sulfur.dioxide,free.sulfur.dioxide,rating)),
        title = "Red Wine Dataset Correlation Matrix", 
        mapping = aes(color = quality), 
        lower = list(continuous = wrap("points", shape = I('.'))),
        upper = list(combo = wrap("box", outlier.shape = I('.'))))

```


Lets first explore what qualities of wine correlate with quality. From the 
above correlation matrix, we see that quality has the highest correlation with 
Alcohol and negative correlation with volatile acidity. 

```{r echo=FALSE, Quality_vs_alcohol}
#Plot relationship between quality and alcohol
ggplot(data=redwine, aes(y=alcohol, x=quality, fill = quality)) + 
  labs(x = 'quality', y = 'alcohol',
       title = 'Quality vs alcohol') +
  geom_boxplot() + 
  theme(legend.position="none")
```

The box plot of relationship of quality vs alcohol is quite interesting - it 
appears that higher quality wines have higher percent of alcohol content. 

```{r echo=FALSE, Quality_vs_volatile.acidity}
#Plot relationship between quality and volatile.acidity
ggplot(data=redwine, aes(y=volatile.acidity, x=quality, fill = quality)) + 
  labs(x = 'quality', y = 'volatile acidity',
       title = 'Quality vs volatile acidity') +
  geom_boxplot() + 
  theme(legend.position="none")
```

Box plot of quality against volatile acidity shows that normally lower quality 
wines have higher level of volatile acidity and vice versa higher quality wines 
have lower level of volatile acidity. This makes sense because volatile acidity 
is the amount of acetic acid in wine, which at too high of levels can lead to 
an unpleasant, vinegar taste. 

```{r echo=FALSE, Quality_vs_citric.acid}
#Plot relationship between quality and citric.acid
ggplot(data=redwine, aes(y=citric.acid, x=quality, fill = quality)) + 
  labs(x = 'quality', y = 'citric acid',
       title = 'Quality vs citric acid') +
  geom_boxplot() + 
  theme(legend.position="none")
```

There is a slight correlation of Quality vs citric.acid and it appears that 
higher quality wines have slightly higher level of citric acid. This also makes 
sense given that citric acid can add 'freshness' and flavor to wines. 

```{r echo=FALSE, Quality_vs_pH}
#Plot relationship between quality and pH
ggplot(data=redwine, aes(y=pH, x=quality, fill = quality)) + 
  labs(x = 'quality', y = 'pH',
       title = 'Quality vs pH') +
  geom_boxplot() + 
  theme(legend.position="none")
```

pH describes how acidic or basic a wine is on a scale from 0 (very acidic) to 
14 (very basic); most wines are between 3-4 on the pH scale. The above boxplot 
shows that wines in our sample are mainly within the 3-4 range. In general, 
better quality wines tend to have a lower pH scale, with some outliers of course.  

Next lets explore which chemical qualities alcohol and volatile.acidity are 
correlated with.

Alcohol has high correlation with Density, so lets plot that first. 

```{r echo=FALSE,message=FALSE, warning=FALSE,Alcohol_vs_Density}
#Plot relationship between alcohol and density

ggplot(aes(x = alcohol, y = density), data = redwine, binwidth = 100) +
  labs(x = 'alcohol', y = 'density',
       title = 'Alcohol vs density') + 
  scale_x_continuous() +
  geom_point(color = I('#882817')) + 
  geom_smooth() 
```

There is negative correlation between density and alcohol which makes sense 
given that the density of water is close to that of water depending on the 
percent alcohol and sugar content.


Per correlation matrix volatile.acidity has the highest correlation with 
citric.acid, lets plot to see how that relationship looks like. 

```{r echo=FALSE,message=FALSE,warning=FALSE, Citric_Acid_vs_Volatile_Acidity}
#Plot relationship between citric.acid and volatile.acidity

ggplot(aes(x = citric.acid, y = volatile.acidity), data = redwine, 
       binwidth = 100) +
  labs(x = 'citric.acid', y = 'volatile.acidity',
       title = 'Citric Acid vs Volatile Acidity') + 
  scale_x_continuous() +
  geom_point(color = I('#882817')) + 
  geom_smooth()
```

There is slightly negative relationship between volatile.acidity and 
citric.acid. The higher levels of citric.acid is associated with lower level of 
volatile.acidity. 


```{r echo=FALSE,message=FALSE, warning=FALSE,Fixed.acidity_vs_density}
#Plot relationship between fixed.acidity and density

ggplot(aes(x = fixed.acidity, y = density), data = redwine, binwidth = 100) +
  labs(x = 'fixed.acidity', y = 'density',
       title = 'Fixed acidity vs density') + 
  scale_x_continuous() +
  geom_point(color = I('#882817')) + 
  geom_smooth()
```

Per above the higher the level of fixed.acidity the higher is the level of 
density. 


```{r echo=FALSE, message=FALSE, warning=FALSE, Citric.acid_vs_Density}
#Plot relationship between citric.acid and density

ggplot(aes(x = citric.acid, y = density), data = redwine, binwidth = 100) +
  labs(x = 'citric.acid', y = 'density',
       title = 'Citric acid vs Density') + 
  scale_x_continuous() +
  geom_point(color = I('#882817')) + 
  geom_smooth()
```

Citric.acid is positively correlated with density level. Higher citric.acid 
indicates higher density. 


```{r echo=FALSE, message=FALSE, warning=FALSE, Citric.acid_vs_Fixed.acidity}
#Plot relationship between fixed.acidity and citric.acid

ggplot(aes(x = fixed.acidity, y = citric.acid), data = redwine, binwidth = 100) +
  labs(x = 'fixed.acidity', y = 'citric.acid',
       title = 'Citric acid vs Fixed acidity') + 
  scale_x_continuous() +
  geom_point(color = I('#882817')) + 
  geom_smooth()
```

There is positive correlation between fixed.acidity and citric.acid levels. 

```{r echo=FALSE, message=FALSE, warning=FALSE, pH_vs_Fixed.acidity}
#Plot relationship between fixed.acidity and pH

ggplot(aes(x = fixed.acidity, y = pH), data = redwine) +
  labs(x = 'fixed.acidity', y = 'pH',
       title = 'pH vs Fixed acidity') + 
  scale_x_continuous() +
  geom_point(color = I('#882817')) + 
  geom_smooth()

```

It appears pH and fixed.acidity have high negative correlation. pH describes 
how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very 
basic); most wines are between 3-4 on the pH scale. Per chart above 
wines with lower levels of fixed.acidity are less acidic. 


```{r echo=FALSE, message=FALSE, warning=FALSE, Citric.acid_vs_pH}
#Plot relationship between citric.acid and pH

ggplot(aes(x = citric.acid, y = pH), data = redwine) +
  labs(x = 'citric.acid', y = 'pH',
       title = 'Citric acid vs pH') + 
  scale_x_continuous() +
  geom_point(color = I('#882817')) + 
  geom_smooth()

```

Per chart higher level of citric.acid are associated with lower pH, 
i.e. higher citric.acid means less acidic. 



# Bivariate Analysis


### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

My main feature of interest in the above analysis was what variables were 
associated with quality of wine. I noticed that quality is has positive
correlation with alcohol and negative correlation with volatile acidity. \

1. It appears that higher quality wines have higher percent of alcohol content.\

2. Box plot of quality against volatile acidity shows that lower quality 
wines have higher level of volatile acidity and higher quality wines 
have lower level of volatile acidity. This makes sense because volatile acidity 
is the amount of acetic acid in wine, which at too high of levels can lead to 
an unpleasant, vinegar taste. \

3. There is a slight correlation of Quality vs citric.acid and it appears that 
higher quality wines have slightly higher level of citric acid. This also makes 
sense given that citric acid can add 'freshness' and flavor to wines. \

4. pH vs Quality shows that wines in our sample are mainly within the 3-4 range. 
In general, better quality wines tend to have a lower pH scale, with some 
outliers of course. \


### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

In my analysis I explored relationships for those variables which appeared to 
have the highest correlation indices. There were some more obvious observations 
like negative correlation between density and alcohol and negative relationship 
between volatile.acidity and citric.acid. \
\
However, the relationship I found interesting was between fixed.acidity and 
other variables. There is positive correlation between fixed.acidity and 
citric.acid levels and citric.acid known to bring freshness and flavor to wine. 
While pH and fixed.acidity have high negative correlation.
Wines with lower levels of fixed.acidity appear less acidic. Also, higher the 
level of fixed.acidity the higher is the level of density. 

### What was the strongest relationship you found?

Fixed.acidity and pH is the strongest relationship I found. 


# Multivariate Plots Section


```{r echo=FALSE, Alcohol_vs_citric.acid_by_quality}

#Plot relationship between alcohol, citric.acid and quality
ggplot(data=redwine, aes(x=alcohol, y=citric.acid, color = quality)) + 
  labs(x = 'alcohol', y = 'citric.acid',
       title = 'Alcohol and Citric.acid by quality') +
  geom_point() +
  scale_color_brewer(type = "div", palette = "RdYlBu", name="Quality", direction=-1)
```

This is interesting. We can see that high alcohol content and high citric.acid 
provide high quality wine, while the low alcohol content and low citric.acid 
get the low quality score. Interestingly, citric.acid impacts quality score 
more than the alcohol content. We can observe this for some high quality wines 
which have lower alcohol content but high citric.acid. 


```{r echo=FALSE, Alcohol_vs_volatile.acidity_by_quality}

#Plot relationship between alcohol, volatile.acidity and quality
ggplot(data=redwine, aes(x=alcohol, y=volatile.acidity, color = quality)) + 
  labs(x = 'alcohol', y = 'volatile.acidity',
       title = 'Alcohol vs Volatile.acidity by quality') +
  geom_point() 
```

Plotting alcohol against volatile.acidity (the two variables highly correlated 
with quality), we see the expected tendency: the higher alcohol and low 
volatile.acidity results in higher quality. However, we see some outliers 
where high alcohol content and higher volatile.acidity results in lower quality 
score vs lower alcohol content and lower volatile.acidity. 


Lets now plot citric.acid_vs_volatile.acidity and split by quality and see what
happens. 

```{r echo=FALSE, citric.acid_vs_volatile.acidity_by_quality}
#Plot relationship between citric.acid, volatile.acidity and quality
ggplot(data=redwine, aes(x=citric.acid, y=volatile.acidity, color = quality)) + 
  labs(x = 'citric.acid', y = 'volatile.acidity',
       title = 'Citric.acid vs Volatile.acidity by quality') +
  geom_point()
```

What we see here is that low citric.acid and high volatile.acidity results in
low quality score, which is expected based on our observations above. 
However, wine with medium volatile.acidity and high citric.acid will still get a
low quality score. This shows that although both important it is actually 
more impactful to have lower volatile.acidity than having high citric.acid.


```{r echo=FALSE, message=FALSE, warning=FALSE, Citric.acid_vs_Fixed.acidity_by_quality}
#Plot relationship between alcohol, fixed.acidity and quality

ggplot(aes(x = citric.acid, y = fixed.acidity), 
       data = redwine) +
  labs(x = 'citric.acid',
       y = 'fixed.acidity', title = 'Citric.acid vs Fixed.acidity by quality') +
  geom_point(aes(color = quality), position = 'jitter') +
  geom_smooth() +
  facet_wrap(~rating) + 
  scale_color_brewer(type = "div", palette = "RdYlBu", name="Quality", direction=-1)
```

Per above chart, we see that positive correlation between fixed.acidity and 
citric.acid is consistent across all three categories of wine (bad, good, 
excellent). However, this relationship is more obvious with highest scoring 
wine category where fixed.acidity level increases along with citric.acid level 
increase. 

```{r echo=FALSE,message=FALSE,warning=FALSE,Density_vs_fixed.acidity_by_rating}
#Plot relationship between fixed.acidity, density and quality

ggplot(aes(x = fixed.acidity, y = density), 
       data = redwine) +
  labs(x = 'fixed.acidity',
       y = 'density', 
       title = 'Density vs. fixed.acidity by rating category') +
  geom_point(aes(color = rating), position = 'jitter', alpha = 0.5) +
  geom_smooth(aes(color = rating)) + 
  scale_color_brewer(type = 'seq') 
```

Fixed.acidity and density relationship by wine category reveals that higher 
quality wines appear less dense compared to other categories for the same level 
of fixed.acidity. 


# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

I explored further some of the relationships explored in previous sections by 
adding quality varibles and seeing if the relationship holds true across all
categories of wine. 

1. Fixed.acidity and citric.acid is consistent 
across all three categories of wine (bad, good, excellent). \
\
2. Alcohol against volatile.acidity show the expected tendency: the higher 
alcohol and low volatile.acidity results in higher quality. \
\
3. As previously noted high alcohol content and high citric.acid 
provide highest quality wine, while the low alcohol content and low citric.acid 
get the lowest quality score. \
\
4. Fixed.acidity and density relationship by wine category reveals that higher 
quality wines appear less dense compared to other categories for the same level 
of fixed.acidity. \

### Were there any interesting or surprising interactions between features?

Some interesting observations are: 

1.Low citric.acid and high volatile.acidity results in
lowest quality score, which is expected based on previous observations. 
However, wine with high volatile.acidity and high citric.acid will still get a
low quality score (4). This shows that although both important it is actually 
more impactful to have lower volatile.acidity than having high citric.acid.\
\
2. In citric.acid vs alcohol chart we notice that citric.acid impacts quality 
score more than the alcohol content. This we can see for quality score 5 which
has higher alcohol content but less citric.acid vs quality score 7 which has
lower alcohol content but higher citric.acid. \
\
3. Alcohol vs volatile.acidity relationship related to quality #5 has high 
alcohol content and higher volatile.acidity results in lower quality score vs 
lower alcohol content and lower volatile.acidity. This means that 
volatile.acidity impacts quality score more than the alcohol content.\


### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

N/A


# Final Plots and Summary

One of the highest correlations we have noted in the dataset is alcohol vs 
quality. 

```{r echo=FALSE, Plot_One}
#plot relationship between alcohol and quality

ggplot(data=redwine, aes(y=alcohol, x=quality, fill = quality)) + 
  labs(x = 'Quality', y = '% of Alcohol content',
       title = 'Quality vs Alcohol') +
  geom_boxplot() + 
  theme(legend.position="none")
```

### Description One

It appears that higher quality wines have higher percent of alcohol content.


### Plot Two

```{r echo=FALSE, Plot_Two}
#Plot relationship between citric.acid, volatile.acidity and quality

ggplot(data=redwine, aes(x=citric.acid, y=volatile.acidity, color = quality)) + 
  labs(x = 'Citric Acid (g/L)', y = 'Volatile Acidity (g/L)',
       title = 'Citric Acid vs Volatile Acidity by Quality') +
  geom_point()
```

### Description Two

Bivariate analysis performed above also showed that there is high positive 
correlation between quality and citric acid and negative correlation between 
quality and volatile acidity. Here I am plotting citric acid against volatile 
acidity by quality score. We can observe that low citric.acid and high 
volatile.acidity results in lowest quality score (3). However, wine with high 
volatile.acidity and high citric.acid will still get a low quality score (4). 
This shows that although both important it is actually 
more impactful to have lower volatile.acidity than having high citric.acid.


### Plot Three

```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}

ggplot(aes(x = fixed.acidity, y = density), 
       data = redwine) +
  labs(x = 'Fixed Acidity (g/L)',
       y = 'Density (g/mL)', 
       title = 'Density vs. Fixed Acidity by Rating Category') +
  geom_point(aes(color = rating), position = 'jitter', alpha = 0.5) +
  geom_smooth(aes(color = rating)) + 
  scale_color_brewer(type = 'seq') 
```


### Description Three

Our bivariate analysis show that there is negative correlation between density 
and alcohol which makes sense given that the density of water is close to that 
of water depending on the percent alcohol and sugar content.We also observe 
that there is positive correlation between fixed.acidity and citric.acid levels. \
\
Therefore, I also want to see how the relationship between fixed.acidity and 
density relationship. Fixed.acidity and density relationship by wine category
reveals that higher quality wines appear less dense compared to other categories
for the same level of fixed.acidity. \

------

# Reflection

There are 1,599 observations in the dataset with 12 variables. At least 3 wine
experts rated the quality of each wine, providing a rating between 0 (very bad)
and 10 (very excellent). \
\
In univariate analysis I noted that our dataset shows that majority of wine are 
ranked in quality score of 5 and 6.\
\
Comparing qualities of bad, good and excellent wines, volatile.acidity and 
citric.acid differs significantly for each rating and therefore hinting those 
are the qualities impacting the quality of wine. Another characteristic that 
would be interesting to explore is the level of alcohol and its impact on 
quality of wine. \
\
Further, I explored bivariate relationships between different chemical 
components. In my analysis I explored relationships for those variables which 
appeared to have the highest correlation indices. There were some more obvious 
observations like negative correlation between density and alcohol and negative 
relationship between volatile.acidity and citric.acid. \
\
There is positive correlation between fixed.acidity and 
citric.acid levels and citric.acid known to bring freshness and flavor to wine. 
While pH and fixed.acidity have high negative correlation.
Wines with lower levels of fixed.acidity appear less acidic. Also, higher the 
level of fixed.acidity the higher is the level of density. \
\
In multivariate analysis we observed that fixed.acidity and citric.acid 
relationship is consistent across all three categories of wine 
(bad,good,excellent). \
\
Also, the higher alcohol and low volatile.acidity results in higher quality. 
High alcohol content and high citric.acid provide highest quality wine, while 
the low alcohol content and low citric.acid get the lowest quality score. 
Fixed.acidity and density relationship by wine category reveals that higher 
quality wines appear less dense compared to other categories for the same level 
of fixed.acidity. \
\
You can find these and other observations described in the above sections. \
\
What would be interesting to add in the dataset for the future is the region 
where wine is from. 




